---
title: "Pteropod biomarker analysis"
author: ""
output: 
  html_document:
    code_folding: hide
    css: kable.css
self_contained: yes
runtime: shiny
---

```{r, message = F, warning = F}
library(knitr)
opts_chunk$set(warning = FALSE, message = FALSE, dev.args = list(bg = 'transparent'), eval = T)

library(tidyverse)
library(vegan)
library(ggord)
library(scales)
library(effects)
library(shiny)
library(plotly)
library(GGally)
library(grid)

source("R/funcs.R")

data(envdat)
data(ptedat)
data(biomod)

envchr <- c('pCO2', 'pH', 'Ara', 'O2', 'Temp', 'Fluor')
ptechr <- c('CAT', 'GR', 'GSHonGSSG', 'GST', 'LPX', 'ORAC', 'SOD', 'ORACvLPX')
```

Pairs plot
```{r, fig.height = 8, fig.width = 8}
dat <- ptedat %>% 
  select(one_of('CTD', ptechr)) %>% 
  na.omit %>% 
  inner_join(envdat, by = 'CTD') %>% 
  select(one_of(c('CTD', envchr, ptechr)))
 
pairs(dat[, -1], gap = 0, cex = 1, col = 'grey')
```

RDA triplot of biomarkers and environmental data
```{r, fig.height = 4, fig.width = 5, eval = T}
dat_frm <- dat %>% 
  data.frame %>% 
  remove_rownames %>% 
  column_to_rownames('CTD')

env <- select(dat_frm, one_of(envchr)) %>% 
  decostand(method = 'range')
pte <- select(dat_frm, one_of(ptechr)) %>% 
  decostand(method = 'range')

mod <- rda(pte, env)
ggord(mod, ptslab = T, parse = T)
```

# Correlations

```{r}
# physiology rda model
dat_cor <- ptedat %>% 
  select(one_of('CTD', ptechr)) %>% 
  # na.omit %>% 
  left_join(envdat, by = 'CTD') %>% 
  data.frame %>% 
  remove_rownames %>% 
  column_to_rownames('CTD')

# all correlations
crs <- crossing(var1 = names(dat_cor), var2 = names(dat_cor)) %>% 
  filter(var1 != var2) %>% 
  rownames_to_column() %>% 
  group_by(rowname) %>% 
  nest %>% 
  mutate(
    crs = map(data, function(x){
      
      # variables
      vr1 <- dat_cor[[x$var1]]
      vr2 <- dat_cor[[x$var2]]
      
      # pearson
      pr_ts <- cor.test(vr1, vr2, method = 'pearson')
      pr_cr <- round(pr_ts$estimate, 2)
      pr_pv <- p_ast(pr_ts$p.value)
      pr <- paste(pr_cr, pr_pv)
    
      out <- data.frame(pr = pr, stringsAsFactors = F)
      return(out)
      
    })
  ) %>% 
  unnest %>% 
  select(-rowname)
```

```{r, fig.height = 6, fig.width = 6}
levs <- c(sort(envchr), sort(ptechr))
labs <- c('Omega[ar]', 'chla', 'O[2]', 'pCO[2]', 'pH', 'Temp', 'CAT', 'GR', 'GSHonGSSG', 'GST', 'LPX', 'ORAC', 'SOD', 'ORACvLPX')
prplo <- crs %>% 
  separate(pr, c('cor', 'sig'), sep = ' ') %>% 
  filter(var1 %in% levs & var2 %in% levs) %>%  
  mutate(
    cor = as.numeric(cor), 
    var1 = factor(var1, levels = rev(levs), labels = rev(labs)), 
    var2 = factor(var2, levels = rev(levs), labels = rev(labs)), 
    sig = gsub('ns', '', sig)
  )

pbase <- theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(), 
  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8), 
  axis.text.y = element_text(size = 8),
  legend.position = c(0.5, 1.2),
  legend.direction = 'horizontal',
  plot.margin = unit(c(8,4,0,0), "lines"),
  strip.background = element_blank(), 
  strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5), 
  panel.background = element_rect(fill = 'black')
  ) 

outlab <- data.frame(
  y = c(4.5, 11.5), 
  lab = c('Cellular', 'Environment')
)

p <- ggplot(prplo) + 
  geom_tile(aes(y = var1, x = var2, fill = cor), colour = 'black') + 
  geom_text(aes(y = var1, x = var2, label = sig)) +
  annotation_custom(grob = textGrob(label = outlab$lab[1], hjust = 0, gp = gpar(cex = 0.7)),
                    ymin = outlab$y[1], ymax = outlab$y[1], xmin = 15, xmax = 15) +
  annotation_custom(grob = textGrob(label = outlab$lab[2], hjust = 0, gp = gpar(cex = 0.7)),
                    ymin = outlab$y[2], ymax = outlab$y[2], xmin = 15, xmax = 15) +
  annotation_custom(grob = textGrob(label = outlab$lab[1], hjust = 0.5, gp = gpar(cex = 0.7)),
                    xmin = outlab$y[1], xmax = outlab$y[1], ymin = 15.5, ymax = 15.5) +
  annotation_custom(grob = textGrob(label = outlab$lab[2], hjust = 0.5, gp = gpar(cex = 0.7)),
                    xmin = outlab$y[2], xmax = outlab$y[2], ymin = 15.5, ymax = 15.5) +
  pbase +
  scale_y_discrete('', expand = c(0, 0), labels = parse(text = rev(labs))) + 
  scale_x_discrete('', expand = c(0, 0), labels = parse(text = rev(labs))) +
  scale_fill_gradient2('Correlation', low = muted("blue"), mid = "white", high = muted("red"), midpoint = 0) +
  geom_hline(yintercept = 8.5, size = 1.5) +
  geom_vline(xintercept = 8.5, size = 1.5) 

# Code to override clipping
gt <- ggplot_gtable(ggplot_build(p))
gt$layout$clip[gt$layout$name == "panel"] <- "off"
grid.draw(gt)
```

# Pairwise VIF among environmental variables

```{r}
vifv <- crossing(var1 = envchr, var2 = envchr) %>% 
  filter(var1 != var2) %>% 
  rownames_to_column() %>% 
  group_by(rowname) %>% 
  nest %>% 
  mutate(
    vifv = map(data, function(x){
      
    vr1 <- x[['var1']]
    vr2 <- x[['var2']]
    frm <- formula(paste(vr1, '~', vr2))
    mod <- lm(frm, dat_frm)
    rsq <- mod %>% 
      summary %>% 
      .$r.squared
    vif <- 1 / (1 - rsq)
    
    return(vif)
    
    })
  ) %>% 
  unnest %>% 
  select(-rowname) %>% 
  mutate(vifv = round(vifv, 2)) %>% 
  spread(var2, vifv) %>% 
  data.frame(stringsAsFactors = F) %>% 
  remove_rownames %>% 
  column_to_rownames('var1') 
```

<div class="verysmall">

```{r}
knitr::kable(vifv, caption = 'VIF')
```

</div>

# Model results {.tabset}

```{r}
# model data
mods <- biomod %>% 
  unnest(data)
mod_all <- biomod
```

## LPX
```{r}
knitr::kable(filter(mods, pte_lab %in% 'LPX'), caption = 'LPX')
```

## ORAC
```{r}
knitr::kable(filter(mods, pte_lab %in% 'ORAC'), caption = 'ORAC')
```

## ORACvLPX
```{r}
knitr::kable(filter(mods, pte_lab %in% 'ORACvLPX'), caption = 'ORACvLPX')
```

## SOD
```{r}
knitr::kable(filter(mods, pte_lab %in% 'SOD'), caption = 'SOD')
```

# {.tabset}

Pairwise models with interactions for relevant response measures, only pairwise models with VIF < 11 are shown

```{r}

selectInput("mod", "Choose model:", mod_all[['Model']])

md <- reactive({

  md <- filter(mod_all, Model %in% input$mod) %>% 
  .$Modobj %>% 
  .[[1]]
  
  return(md)
  
})

trm <- reactive({
  
  coefnms <- md() %>% coefficients %>% names
  if(length(coefnms) < 3) return(NULL)
  else out <- coefnms[3]
  
  if(input$cvr) out <- coefnms[2]
  
  return(out)

})

prd_vl <- reactive({
  
  x <- md()$model %>% 
    .[, -1] %>% 
    data.frame %>% 
    as.list %>% 
    map(range) %>%
    map(function(x) seq(x[1], x[2], length = 100))
  
  nms <- names(x) 
  x <- crossing(x[[1]], x[[2]])
  names(x) <- nms
  
  prd_vl <- predict(md(), newdata = x) %>% 
    data.frame(., x)
  names(prd_vl)[1] <- all.vars(formula(md()))[1]

  return(prd_vl)
  
})

renderPrint({summary(md())})

renderPlot({
  toplo <- md()$model %>% 
    gather('var', 'val', -1)

  ggplot(toplo, aes_string(x='val', y=names(toplo)[1])) +
    geom_point() + 
    facet_wrap(~var, ncol = 2, scales = 'free_x') + 
    theme_bw(base_family = 'serif') + 
    geom_smooth(method = 'lm') + 
    theme(axis.title.x = element_blank())

}, height = 300, width = 600)

selectInput("cvr", "Switch covariate:", c(F, T))

renderPlot({
  
  validate(
    need(!is.null(trm()), 'Cannot show plots for models with one predictor')
  )
  
  # colors
  cols <- RColorBrewer::brewer.pal(9, 'RdBu')
  
  # get data to plot
  pl <- md() %>% 
    get_pldat(., trm())
  pl1 <- pl[[1]]
  pl2 <- pl[[2]]
  
# names for aes_string
nms <- names(pl1)

p1 <- ggplot() +
  geom_ribbon(data = pl1, aes_string(x = nms[3], ymin = 'lo', ymax = 'hi', group = nms[4]), alpha = 0.5, fill = 'grey') +
  geom_line(data = pl1, aes_string(x = nms[3], y = nms[1], group = nms[4], colour = nms[4]), size = 1) + 
  geom_text(data = pl2, aes(x= x, y = y, label = lab), hjust = 0) +
  theme_bw() +
  scale_x_continuous(nms[3]) +
  scale_colour_gradientn(colours = cols) + 
  scale_y_continuous(nms[1])

return(p1)
  
}, height = 300, width = 600)

renderPlotly({
  
  # observed data
  toplo <- md()$model
  req(ncol(toplo) >= 3)
  
  names(toplo) <- c('z', 'x', 'y')
  
  # surface data
  dat <- prd_vl() %>% 
    spread_(key_col = names(.)[2], value_col = names(.)[1]) %>% 
    as.matrix
  x <- colnames(dat)[-1] %>% as.numeric
  y <- dat[, 1]
  dat <- dat[, -1]
  
  scene <- list(
    aspectmode = 'manual', 
    aspectratio = list(x = 1, y = 1, z = 0.6), 
    xaxis = list(title = names(prd_vl()[2])),
    yaxis = list(title = names(prd_vl()[3])),
    zaxis = list(title = names(prd_vl()[1]))
  )
  
  p <- plot_ly() %>% 
    add_surface(z = ~dat, x = ~x, y = ~y, colors = rev(RColorBrewer::brewer.pal(11, 'Spectral'))) %>% 
  add_markers(data = toplo, x = ~x, y = ~y, z = ~z) %>% 
    layout(scene = scene, width = 800, height = 400)
  p
})
```
